---
title: "FitnessData"
output: html_document
---

#Introduction
Analysis of fitness data from fitness trackers like fitbit over 2 months.
Data is fed for 2 months at a time interval of 5 mins each. There are some empty logs as well.

#Reading the file from the link provided and loading it on to be read.

```{r}
library(data.table)
library(ggplot2)

fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = paste0(getwd(), '/repdata%2Fdata%2Factivity.zip'), method = "curl")
unzip("repdata%2Fdata%2Factivity.zip",exdir = "data")

dt<- data.table::fread(input= 'data/activity.csv')
```

#Histogram of total number of steps taken across all the days
```{r}
Totalsteps<- dt[, c(lapply(.SD, sum, na.rm= FALSE)), .SDcols= c("steps"), by= .(date)]
head(Totalsteps, 10)

ggplot(Totalsteps, aes(x=steps)) +
  geom_histogram(fill= 'blue', binwidth = 750)+
  labs( title= 'Total Number of Steps', x='Steps', y='Freq')
```

#Mean and Median of data
```{r}
Totalsteps[, .(MeanS= mean(steps, na.rm= TRUE), MedianS= median(steps, na.rm= TRUE))]
```

#Time Series Plot

```{r}
interval<- dt[, c(lapply(.SD, mean, na.rm= TRUE)), .SDcols= c('steps'), by= .(interval)]
ggplot(interval, aes(x=interval, y=steps))+
  geom_line(color= 'red', size=0.75)+
  labs(title= 'average steps over the course of the day', x='Time', y='Steps')
```

#Locating and Replcing missing values

Step1- Finding number of missing values.
```{r}
dt[is.na(steps), .N]
```
Step2- Replacing them with median of complete dataset
```{r}
dt[is.na(steps), 'steps']<- dt[, c(lapply(.SD, median, na.rm= TRUE)), .SDcols= c('steps')]
data.table::fwrite(x= dt, file= 'data/newactivity.csv', quote = FALSE)
```

#Calculating new mean and median

```{r}
TotalStepsNew <- dt[, c(lapply(.SD, sum, na.rm= FALSE)), .SDcols= c('steps'), by= .(date)]
TotalStepsNew[, .(MeanNew= mean(steps, na.rm= FALSE), MedianNew= median(steps, na.rm= FALSE))]
```

Hence, mean and median both cahnge after injecting values in place of null characters.

#New Histogram

```{r}
TotalStepsNew<- dt[, c(lapply(.SD, sum, na.rm= FALSE)), .SDcols= c("steps"), by= .(date)]
head(Totalsteps, 10)

ggplot(TotalStepsNew, aes(x=steps)) +
  geom_histogram(fill= 'blue', binwidth = 750)+
  labs( title= 'Total Number of Steps', x='Steps', y='Freq')
```

#Seperating on basis of weekend and weekdays

```{r}
dt[, date := as.POSIXct(date, format = "%Y-%m-%d")]
dt[, `Day of Week`:= weekdays(x = date)]
dt[grepl(pattern = "Monday|Tuesday|Wednesday|Thursday|Friday", x = `Day of Week`), "weekday or weekend"] <- "weekday"
dt[grepl(pattern = "Saturday|Sunday", x = `Day of Week`), "weekday or weekend"] <- "weekend"
dt[, `weekday or weekend` := as.factor(`weekday or weekend`)]
head(dt, 10)
```

```{r}
dt[is.na(steps), "steps"] <- dt[, c(lapply(.SD, median, na.rm = TRUE)), .SDcols = c("steps")]
IntervalDT <- dt[, c(lapply(.SD, mean, na.rm = TRUE)), .SDcols = c("steps"), by = .(interval, `weekday or weekend`)] 
ggplot(IntervalDT , aes(x = interval , y = steps, color=`weekday or weekend`)) + geom_line() + labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") + facet_wrap(~`weekday or weekend` , ncol = 1, nrow=2)
```